---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
options(tibble.print_min = 20, tibble.print_max = 20)
```

# multides

<!-- badges: start -->
<!-- badges: end -->

The `multides` package is a companion R package of the project "MULTI-DES: 
Multilevel Design Parameters and Effect Size Benchmarks for Students’ Competencies,"
as well as its follow-up project "MULTI-DES 2: Multilevel Design Parameters 
for Sample Size Planning of Randomized Intervention Studies in Preschool, 
Elementary and Secondary School."

`multides` compiles tools that were developed to facilitate the analyses conducted 
within the MULTI-DES project framework. The functions may be used to replicate 
the R code that accompanies the manuscripts prepared within MULTI-DES. 
All R scripts are shared via the Open Science Framework (OSF; see below). 

Nevertheless, the application scope of `multides` is not limited to the 
analyses specific to MULTI-DES. The functions provided can also be used in other
contexts of (multilevel) data analysis, for instance, to easily generate 
overviews of descriptive statistics, single- and multilevel 
correlation matrices, or--most importantly--to calculate (multilevel) design parameters 
such as intraclass correlation coefficients and explained variances by covariates 
at each hierarchical level with corresponding standard errors, 
based on the variance components estimated from single- or multilevel models.

## The MULTI-DES Project

MULTI-DES is funded by the German Research Foundation (DFG) and aims at investigating 
(1) single- and multilevel design parameters that are needed to efficiently plan
adequately powered individually and cluster randomized trials on various outcomes 
(viz., achievement/cognitive outcomes and socio-emotional learning outcomes) 
in preschool, elementary and secondary school, as well as
(2) effect size benchmarks in terms of academic growth and performance gaps between 
schools or policy-relevant groups to appropriately interpret and communicate the 
results of such studies. 
In MULTI-DES, rich data from several German (longitudinal) large-scale assessments 
were used to apply multilevel and structural equation modeling. For more information 
on MULTI-DES, visit https://www.uni-potsdam.de/en/quantmethoden/forschung.

### Publications and Preprints 

Brunner, M., Stallasch, S. E., Artelt, C., Hedges, L. V., & Lüdtke, O. (2024). 
An individual participant data meta-analysis to support power analyses for 
randomized intervention studies in preschool: Cognitive and socio-emotional 
learning outcomes. PsyArXiv. https://doi.org/10.31234/osf.io/dkw42 <br>
[Preprint](https://osf.io/preprints/psyarxiv/dkw42)

Brunner, M., Stallasch, S. E., & Lüdtke, O. (2023). Empirical benchmarks to 
interpret intervention effects on student achievement in elementary and 
secondary school: Meta-analytic results from Germany. 
*Journal of Research on Educational Effectiveness*, *17*(1), 1–39. 
https://doi.org/10.1080/19345747.2023.2175753 <br>
[R code](https://osf.io/wp5rf)

Stallasch, S. E., Lüdtke, O., Artelt, C., & Brunner, M. (2021). 
Multilevel design parameters to plan cluster-randomized intervention studies on 
student achievement in elementary and secondary school. 
*Journal of Research on Educational Effectiveness*, *14*, 172–206. 
https://doi.org/10.1080/19345747.2020.1823539 <br> 
[R code](https://osf.io/2w8nt)  

Stallasch, S. E., Lüdtke, O., Artelt, C., Hedges, L. V., & Brunner, M. (2024). 
Single- and multilevel perspectives on covariate selection in randomized 
intervention studies on student achievement. *Educational Psychology Review*, 
*36*, 112. https://doi.org/10.1007/s10648-024-09898-7 <br>
[R code](https://osf.io/nhx4w)




## Installation

You can install the development version of `multides` from [GitHub](https://github.com/) with:

``` r
# install.packages("devtools")
devtools::install_github("sophiestallasch/multides")
```

## Usage

```{r example}
library(multides)

# calculate a range of descriptive statistics for a series of numeric variables,
# grouped by school type
describe_stats(studach, gender:read, ts_name)

# calculate multilevel correlations between group means at the school level
correlate_ml(studach, gender:read, id_sch)
```


## Help & Suggestions

If you encounter a bug, have questions or suggestions for improvement, please file an issue on 
[GitHub](https://github.com/sophiestallasch/multides/issues) or [email me](mailto:sophie.stallasch@gmail.com), 
possibly including a minimal reproducible example.
