---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
options(tibble.print_min = 20, tibble.print_max = 20)
```

# multides

<!-- badges: start -->
<!-- badges: end -->

The `multides` package is a companion `R` package of the project 'MULTI-DES: 
Multilevel Design Parameters and Effect Size Benchmarks for Students’ Competencies'.

`multides` compiles tools that were developed for the analyses conducted 
within the 'MULTI-DES' project framework. The functions are used to replicate 
the `R` code that accompanies the manuscripts prepared within 'MULTI-DES'.
The `R` scripts are shared via the Open Science Framework (OSF; see below). 

Nevertheless, the application scope of `multides` is not limited to the 
analyses specific to 'MULTI-DES'. The functions provided can also be used in other
contexts of (multilevel) data analysis, for instance, 
to easily generate overviews of descriptive statistics, single- and multilevel 
correlation matrices, or to calculate (multilevel) design parameters such as 
intraclass correlation coefficients and explained variances by covariates 
at each hierarchical level with corresponding standard errors, 
based on the variance components estimated from single- or multilevel models.

## The 'MULTI-DES' Project

'MULTI-DES' is funded by the German Research Foundation (DFG) and aims at investigating 
(1) single- and multilevel design parameters that are needed to efficiently plan
adequately powered individually and cluster-randomized trials on student achievement, 
respectively, as well as
(2) effect size benchmarks in terms of academic growth and performance gaps between schools
to appropriately interpret and communicate the results of such studies. 
In 'MULTI-DES', representative data from three German longitudinal large-scale assessments 
were used to apply multilevel and structural equation modeling. For more information on 
'MULTI-DES', visit https://www.uni-potsdam.de/en/quantmethoden/forschung.

The 'MULTI-DES' project contains 3 manuscripts so far (that are either published or in progress):  
1) Stallasch, S. E., Lüdtke, O., Artelt, C., & Brunner, M. (2021). 
Multilevel design parameters to plan cluster-randomized intervention studies on student achievement 
in elementary and secondary school. Journal of Research on Educational Effectiveness, 14, 172–206. 
https://doi.org/10.1080/19345747.2020.1823539  
The `R` code for this manuscript is accessible via the OSF under https://osf.io/2w8nt.  

2) Stallasch, S. E., Lüdtke, O., Artelt, C., & Brunner, M. (in progress). 
Single- and multilevel perspectives on covariate selection in randomized trials on student achievement.  

3) Brunner, M., Stallasch, S. E., Lüdtke, O., & Artelt, C. (in progress). 
Empirical benchmarks to interpret effect sizes related to K-12 students’ achievement: 
An update with results from Germany.

## Installation

You can install the development version of `multides` from [GitHub](https://github.com/) with:

``` r
# install.packages("devtools")
devtools::install_github("sophiestallasch/multides")
```

## Usage

```{r example}
library(multides)

# calculate a range of descriptive statistics for a series of numeric variables,
# grouped by school type
describe_stats(studach, gender:read, ts_name)

# calculate multilevel correlations between group means at the school level
correlate_ml(studach, gender:read, id_sch)
```


## Help & Suggestions

If you encounter a bug, have questions or suggestions for improvement, please file an issue on 
[GitHub](https://github.com/sophiestallasch/multides/issues) or [email me](mailto:sophie.stallasch@gmail.com), 
possibly including a minimal reproducible example.
